---
title: "BLCMs with Covariates"
author:
- Sonja Hartnack
- Valerie Hungerbühler
- Eleftherios Meletis
date: '2022-07-14'
output:
  beamer_presentation: default
  slidy_presentation: default
  ioslides_presentation: default
params:
  presentation: yes
subtitle: CA18208 HARMONY Zurich Training School - https://harmony-net.eu/
---

```{r rendering, eval=FALSE, include=FALSE}
# To render this as PDF (beamer) slides run:
rmarkdown::render('Day_2_Session_3.Rmd', 'beamer_presentation', params=list(presentation=TRUE))
```

```{r setup, include=FALSE}
library("tidyverse")
library("runjags")
library("rjags")
runjags.options(silent.jags=TRUE, silent.runjags=TRUE)
set.seed(2022-07-14)

# Reduce the width of R code output for PDF only:
if(params$presentation){
  knitr::knit_hooks$set(size = function(before, options, envir) {
    if(before){
      knitr::asis_output(paste0("\\", options$size))
    }else{
      knitr::asis_output("\\normalsize")
    }
  })
  knitr::opts_chunk$set(size = "scriptsize")
}

# Collapse successive chunks:
space_collapse <- function(x){ gsub("```\n*```r*\n*", "", x) }
# Reduce space between chunks:
space_reduce <- function(x){ gsub("```\n+```\n", "", x) }
knitr::knit_hooks$set(document = space_collapse)


# To collect temporary filenames:
cleanup <- character(0)
```


## Covariates: Latent variable logistic regression (Lewis 2012)
- A binomial regression model with a logit link function between the latent true prevalence and covariates associated with disease occurrence can be definded as follows, for covariate pattern $i$
$Pr(Y_i= y_i|n_i)=\binom{n_i}{k_i}q_i^{y_i}(1-q_i)^{n_i-y_i}$
where $q_i=Se\pi_i+(1-Sp)(1-\pi_i)$ and $log(\frac{\pi_i}{1-\pi_i})=x_i^T\beta$
- When $Se=1$ and $Sp=1$ then the model
reverts to the classical logistic regression model

## Ex 8 Covariates 
- Explore the data set 'echinococcus.xlsx'
    PCR for either E. multilocularis or E. granulosus, ELISA for both, eggs found by arecoline purgation, Taenia co-infection, age and sex 
- Run classical 'risk factor analysis': is sex, Taenia co-infection or age a risk factor for echinococcus (PCR-prevalence, seroprevalence or purges)? Obtain p-values and ORs with confidence intervals. 


## Ex 9 Covariates 
- Prepare the data set in the correct format (dump, add ones) for BLCM
- Run a model for three tests (assume a very high sensitivity for arecoline purgation)
- Try different priors
- Evidence of conditional dependencies?
- obtain DICs
  'runjags.version.deviance.R'
- Run models with covariates included 
    - 'runjags_version_cestode_update.R','model.cestode.bug', 'model_without_cestode.bug'
- Is there evidence for a covariate effect on the prevalence?
- Compare your findings with Ex.8

## Ex 8 Covariates 
- Explore the data set 'echinococcus.xlsx'
    PCR for either E. multilocularis or E. granulosus, ELISA for both, eggs found by arecoline purgation, Taenia co-infection, age and sex 
- Run classical 'risk factor analysis': is sex, Taenia co-infection or age a risk factor for echinococcus (PCR-prevalence, seroprevalence or purges)? Obtain p-values and ORs with confidence intervals. 


## Ex 9 Covariates 
- Prepare the data set in the correct format (dump, add ones) for BLCM
- Run a model for three tests (assume a very high sensitivity for arecoline purgation)
- Try different priors
- Evidence of conditional dependencies?
- obtain dics
- Run models with covariates included 
    - 'runjags_version_cestode_update.R','model.cestode.bug', 'model_without_cestode.bug'
- Is there evidence for a covariate effect on the prevalence?
- Compare your findings with Ex.8


