---
title: "MCMC - Stats and jAGS"
author:
- Sonja Hartnack
- Valerie Hungerbühler
- Eleftherios Meletis
date: '2022-07-14'
output:
  beamer_presentation: default
  slidy_presentation: default
  ioslides_presentation: default
params:
  presentation: yes
subtitle: CA18208 HARMONY Zurich Training School - https://harmony-net.eu/
---

```{r rendering, eval=FALSE, include=FALSE}
# To render this as PDF (beamer) slides run:
rmarkdown::render('Day_1_Session_3.Rmd', 'beamer_presentation', params=list(presentation=TRUE))
```

```{r setup, include=FALSE}
library("tidyverse")
library("runjags")
library("rjags")
runjags.options(silent.jags=TRUE, silent.runjags=TRUE)
set.seed(2022-07-14)

# Reduce the width of R code output for PDF only:
if(params$presentation){
  knitr::knit_hooks$set(size = function(before, options, envir) {
    if(before){
      knitr::asis_output(paste0("\\", options$size))
    }else{
      knitr::asis_output("\\normalsize")
    }
  })
  knitr::opts_chunk$set(size = "scriptsize")
}

# Collapse successive chunks:
space_collapse <- function(x){ gsub("```\n*```r*\n*", "", x) }
# Reduce space between chunks:
space_reduce <- function(x){ gsub("```\n+```\n", "", x) }
knitr::knit_hooks$set(document = space_collapse)


# To collect temporary filenames:
cleanup <- character(0)
```


## Markov chain
- Definition: A random process that undergoes transitions from one state to another on a state space

https://setosa.io/blog/2014/07/26/markov-chains/

---

## Markov chain

![](figs/markov.pdf)

---

## Markov chain

- Markov property: A Markov chain possesses a property that is characterised as
"memoryless": the probability of the next state depends only on
the current state and not on the sequence of events that preceded
it.

---

- Markov property defined as:

![](figs/mcprop.pdf)

---

## Ergodicity

- This means that during numerous iterations, the chain will explore every point (or possible state) and will do so proportionally to its probability.
- To be considered ergodic, the Markov chain must be
    - irreducible: for every state there is a positive probability of moving to any other state
    - aperiodic: the chain must not get trapped in cycles

---

## Reversibility: reversible chains
- For a stationary distribution $\pi$ and a transition matrix $P$, the reversibility condition can be written as
$\pi(x)P(x,y) = \pi(y)P(y,x)$, for all $(x,y) \in S$

---

## Link to Metropolis-Hastings and Gibbs

In his famous paper from 1953, Metropolis showed how to construct a Markov chain with stationary distribution $\pi$ such that

\[\pi(x)=p_x,x \in S\]

Here the first step to obtaining the stationary distribution of a Markov chain is to prove that the probabilities of a distribution satisfy the reversibility condition.

===

## JAGS

- JAGS uses the BUGS language

  * This is a declarative (non-procedural) language
  * The order of statements does not matter
  * The compiler converts our model syntax into an MCMC algorithm with appropriately defined likelihood and prior
  * You can only define each variable once!!!

. . .

- Different ways to run JAGS from R:

  - rjags, runjags, R2jags, jagsUI

- See http://runjags.sourceforge.net/quickjags.html

---

A simple JAGS model might look like this:

```{r, echo = TRUE}
model_definition <- "model{
  # Likelihood part:
  Positives ~ dbinom(prevalence, TotalTests)
  
  # Prior part:
  prevalence ~ dbeta(2, 2)
  
  # Hooks for automatic integration with R:
  #data# Positives, TotalTests
  #monitor# prevalence
  #inits# prevalence
}
"
cat(model_definition, file='basicjags.bug')
```


---

There are two model statements:

```{r, echo= TRUE, eval=FALSE}
# Likelihood part:
Positives ~ dbinom(prevalence, TotalTests)
```

- states that the number of $Positive$ test samples is Binomially distributed with probability parameter $prevalence$ and total trials $TotalTests$

. . .

```{r, echo= TRUE, eval=FALSE}
# Prior part:
prevalence ~ dbeta(2,2)
```

- states that our prior probability distribution for the parameter $prevalence$ is Beta(2,2)



---

The other lines in this model:

```{r, echo= TRUE, eval=FALSE}
#data# Positives, TotalTests
#monitor# prevalence
#inits# prevalence
```

are automated hooks that are only used by runjags.

---

To run this model, copy/paste the code above into a new text file called "basicjags.bug" in the same folder as your current working directory.  Then run:

```{r, echo= TRUE}
library('runjags')

# data to be retrieved by runjags:
Positives <- 7
TotalTests <- 10

# initial values to be retrieved by runjags:
prevalence <- list(chain1=0.05, chain2=0.95)
```

---

```{r include=FALSE}
runjags.options(silent.jags=TRUE)
```



```{r message=FALSE, warning=FALSE, results='hide'}
results <- run.jags('basicjags.bug', n.chains=2, burnin=5000, sample=10000)
```


First check the plots for convergence:

```{r eval=FALSE, include=TRUE}
plot(results)
```

```{r include=FALSE}
pt <- plot(results)
```

---

Trace plots: the two chains should be stationary:

```{r echo=FALSE}
print(pt[[1]])
```

---

ECDF plots: the two chains should be very close to each other:

```{r echo=FALSE}
print(pt[[2]])
```

---

Histogram of the combined chains should appear smooth:

```{r echo=FALSE}
print(pt[[3]])
```

---

Autocorrelation plot tells you how well behaved the model is:

```{r echo=FALSE}
print(pt[[4]])
```

---

Then check the effective sample size (SSeff) and Gelman-Rubin statistic (psrf):

```{r, include = TRUE}
results
```


Reminder:  we want psrf < 1.05 and SSeff > 1000

