---
title: "STARD-BLCM Guidelines"
author:
- Sonja Hartnack
- Valerie Hungerbühler
- Eleftherios Meletis
date: '2022-07-15'
output:
  beamer_presentation: default
  slidy_presentation: default
  ioslides_presentation: default
params:
  presentation: yes
subtitle: CA18208 HARMONY Zurich Training School - https://harmony-net.eu/
---

```{r rendering, eval=FALSE, include=FALSE}
# To render this as PDF (beamer) slides run:
rmarkdown::render('Day_2_Session_1.Rmd', 'beamer_presentation', params=list(presentation=TRUE))
```

```{r setup, include=FALSE}
set.seed(2022-07-15)

# Reduce the width of R code output for PDF only:
if(params$presentation){
  knitr::knit_hooks$set(size = function(before, options, envir) {
    if(before){
      knitr::asis_output(paste0("\\", options$size))
    }else{
      knitr::asis_output("\\normalsize")
    }
  })
  knitr::opts_chunk$set(size = "scriptsize")
}

# Collapse successive chunks:
space_collapse <- function(x){ gsub("```\n*```r*\n*", "", x) }
# Reduce space between chunks:
space_reduce <- function(x){ gsub("```\n+```\n", "", x) }
knitr::knit_hooks$set(document = space_collapse)


# To collect temporary filenames:
cleanup <- character(0)
```


## STARD-BLCM guidelines

Update of STARD2015 statement/checklist for diagnpstic test evaluation studies that use Bayesian Latent Class models (BLCMs).

Checklist of STARD items for studies that use BLCMs to estimate diagnostic test accuracy in the absence of a reference standard. 


#### Objective: Facilitate improved quality of reporting on the design, conduct and results of diagnostic accuracy studies that use BLCMs.

  * Examples and elaborations for each of the modified items are discussed in Kostoulas et al (2017). 
  
Checklist available as Word document in Session's folder.

--- 

## Group Work

- Please have a look at the paper from Kostoulas et al. (2017) 'STARD-BLCM: Standards for the Reporting of Diagnostic accuracy studies that use Bayesian Latent Class Models'

- Split in groups of five and read one of the available three papers (Bronsvoort (2019), Bodenham  (2021) or Lahuerta (2018)) and discuss if they comply with the STARD-BLCM guidelines. 
    * The papers are located in: Day 2 - Session 1 - STARD-BLCM papers. 

- Prepare a brief summary-presentation (.ppt file), summarising the paper and assessing compliance with STARD-BLCM guidelines.

--- 

## Open Checklist
